---
title: "Get Started with C3NA"
author: "Kuncheng Song"
date: "3/15/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started with C3NA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Step 1. Install R and R Studio
To run the application locally, you are required to install R and R Studio. C3NA was built under the following versions:

+ R Version 4.1.0
  + Please check CRAN (https://cran.r-project.org/) for the installation of R.
  
+ R Studio Version 1.4.1717
  + Please check https://www.rstudio.com/ for the installation of RStudio.


### Step 2. Install from Github
```{r, eval = FALSE}
# install.packages("devtools")
library(devtools)
install_github("zhouLabNCSU/C3NA")
```

### Step 3. Check if all dependent packages are installed properly
Because of the interactive R Shiny environment, C3NA required a large number of packages to be installed. 
Please run the following function to check if all packages are successfully installed. The function will output a table, and please make sure the column under "Found" are all **TRUE**

```{r}
library(C3NA)
C3NAPackageCheck()
```

Once the required package check is complete, you will be able to use C3NA package, and please check out a full tutorials with demo under **Articles** tab on the top guide bar. 

