---
title: "Get Started with C3NA"
author: "Kuncheng Song"
date: "3/15/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started with C3NA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Step 1. Install R and R Studio
To run the application locally, you are required to install R and R Studio. C3NA was built under the following versions:

+ R Version 4.1.0
  + Please check CRAN (https://cran.r-project.org/) for the installation of R.
  
+ R Studio Version 1.4.1717
  + Please check https://www.rstudio.com/ for the installation of RStudio.


### Step 2 Check if all dependent packages are installed properly
```{r}
## CRAN Repo
.cran_packages <- c("stringr", "tibble", "dplyr", "readr", "stats", "visNetwork", 
                    "igraph", "scales", "magrittr", "tidyr", "randomcoloR", 
                    "ggpubr", "ggplot2",  "plotly", "colorspace", 
                    "reshape2", "shiny", "shinyWidgets", "shinydashboard", 
                    "shinyjs", "reactable", "DT", "pheatmap", "dynamicTreeCut", 
                    "cluster", "WGCNA", "RColorBrewer", "base64enc", "htmlwidgets",
                    "foreach")
.inst <- .cran_packages %in% installed.packages()
if (any(!.inst)){
  install.packages(.cran_packages[!.inst],repos = "http://cran.rstudio.com/",
                   dependencies = TRUE)
}

## Bioconductor Repo
.bioc_packages <- c("phyloseq", "qvalue", "impute", 'GO.db', 'preprocessCore')
.inst <- .bioc_packages %in% installed.packages()
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
if(any(!.inst)){
  BiocManager::install(.bioc_packages[!.inst])
}
## GitHub Repo
.git_packages <- c("SpiecEasi", "metagMisc")
.git_packagesPath <- c("zdk123/SpiecEasi", "vmikk/metagMisc")
.inst <- .git_packages %in% installed.packages()
if (any(!.inst)){
  devtools::install_github(.git_packagesPath[!.inst])
}
```
Once the required package check is complete, you will be able to use C3NA package, and please check out a full tutorials with demo under **Articles** tab on the top guide bar. 

### Step 3. Install from Github
```{r, eval = FALSE}
# install.packages("devtools")
library(devtools)
install_github("zhouLabNCSU/C3NA")
```

### Step 4. Package Check
```{r}
library(C3NA)
C3NAPackageCheck()
```
Because of the interactive R Shiny environment, C3NA required a large number of packages to be installed. 
Please run the following function to check if all packages are successfully installed. The function will output a table, and please make sure the column under "Found" are all **TRUE**



