---
title: "Differential Abundance Analysis"
author: "Kuncheng Song"
date: "3/16/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Differential Abundance Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a brief tutorial for conducting differential abundance analysis between two phenotypes and how to correctly add any types of DA analysis to C3NA object for interactive investigation of the co-occurence network. There are three differential abundance analysis done, including ANCOMBC, ALDEx2, and MaAsLin2, all of which conduct differential abundance analysis for each of the taxonomic levels individually. The results are summarized into a single DA table to be added to C3NA. As recommended by other studies, we will filter the count matrices from each taxonomic levels by 10% prevalance prior to running the DA methods. 

## Library
```{r, warning=FALSE}
library(ANCOMBC)
library(ALDEx2)
library(Maaslin2)
library(curl)
library(DT)
library(phyloseq)
library(metagMisc)
suppressPackageStartupMessages(library(tidyverse))
```

## Step 1. Loading the phyloseq data from GitHub
The downloaded data is a phyloseq object with 261 samples, of which 127 are Colorectal Cancer Patients and 134 are healthy control. 
```{r}
githubURL <- ("https://github.com/zhouLabNCSU/C3NA_ScriptsAndData/raw/main/RPackageTutorialData/Post-initiateC3NA/cancer_dada2_CancerAndNormal.rds")
CancerAndNormal <- readRDS(url(githubURL, method="libcurl"))
print(CancerAndNormal)
```

## Step 2. Generating the Data Matrices for Differential Abundance Analysis
```{r}
# Remove OTUs that do not present among these phenotypes and reformat the data
newPS_otu = as.data.frame(as.matrix(otu_table(CancerAndNormal)))
newPS_otu = newPS_otu[which(rowSums(newPS_otu) > 0),]
newPS_tax = as.data.frame(as.matrix(tax_table(CancerAndNormal)))
newPS_tax = newPS_tax[rownames(newPS_otu), ]
newPS_meta = as.data.frame(as.matrix(sample_data(CancerAndNormal)))
newPS_meta = subset(newPS_meta, diagnosis %in% c("Normal", "Cancer"))
newPS_ps = phyloseq(otu_table(newPS_otu, taxa_are_rows=TRUE),
                             tax_table(as.matrix(newPS_tax)),
                             sample_data(newPS_meta))
print(newPS_ps)
DT::datatable(head(newPS_tax))
```
*Ideally, this should match the taxa names used for the C3NA. Ideally, there should not be any t*

```{r}
# Create a phyloseq object for each taxonomic level
taxaLvls <- c("Phylum", "Class", "Order", "Family", "Genus", "Species")
taxaHeaders <- c("p", "c", "o", "f", "g", "s")
taxaCountMatrices_Abs = list()
psList <- list()
for(t in seq_along(taxaLvls)){
  curLevel = taxaLvls[t]
  psTemp = tax_glom(newPS_ps, curLevel, NArm = FALSE)
  curOTUs = as.data.frame(as.matrix(otu_table(psTemp)))
  ## Summing NAs
  curOTUs$TEMP = as.character(as.data.frame(as.matrix(tax_table(psTemp)))[, curLevel])
  curOTUs <- curOTUs %>%
    mutate(TEMP = ifelse(is.na(TEMP), "NA", TEMP)) %>%
    group_by(TEMP) %>%
    summarise_if(is.numeric, sum) %>%
    column_to_rownames("TEMP")
  ## Add the prefix of taxa
  curHeader = taxaHeaders[t]
  rownames(curOTUs) <- paste0(curHeader,"_",rownames(curOTUs))

  ## Phyloseq Object
  curData = curOTUs
  curData[is.na(curData)] <- 0
  curData <- curData[rowSums(curData)>0, ]  
  tempTaxa = matrix(c(rownames(curData)), ncol = 1)
  colnames(tempTaxa) <- "MixedTaxa"
  rownames(tempTaxa) <- tempTaxa[, 1]
  tempPS = phyloseq(otu_table(curData, taxa_are_rows=TRUE),
                    tax_table(tempTaxa),
                    sample_data(psTemp))

  ## Save the Objects
  taxaCountMatrices_Abs[[curLevel]] = curOTUs
  psList[[taxaLvls[t]]] <- tempPS
}
```

#### Create a DA Summary Table
```{r}
DA_Summary = data.frame(
  TaxaName = NULL, DAMethod = NULL, DiffAbn = NULL
)
```

## Step 3. ANCOMBC
```{r}
for(i in seq_along(taxaLvls)){
  print(paste0(taxaLvls[i]))
  curSample = psList[[taxaLvls[i]]]
  tempOriNTaxa = nrow(otu_table(curSample))
  curSample = phyloseq_filter_prevalence(curSample, prev.trh = 0.1, abund.trh = NULL,
  	threshold_condition = "AND", abund.type = "total")
  ## ANCOMBC with BH
  curANCOMBC = ANCOMBC::ancombc(phyloseq = curSample, formula = "diagnosis",
                                p_adj_method = "BH", zero_cut = 1, 
                                lib_cut = 0, group = "diagnosis",
                                struc_zero = TRUE, neg_lb = TRUE, 
                                tol = 1e-5, max_iter = 100, conserve = FALSE, 
                                alpha = 0.05, global = FALSE)
  curANCOMBC_Res = curANCOMBC$res
  DA_Summary_Summary_temp = data.frame(
      DAMethod = "ANCOMBC",
      TaxaName = rownames(curANCOMBC_Res$beta),
      diff_abn = curANCOMBC_Res$diff_abn[,1])
  DA_Summary = rbind(DA_Summary, DA_Summary_Summary_temp)
}
```


## Step 4. ALDEx2
```{r}
for(i in seq_along(taxaLvls)){
  print(paste0(taxaLvls[i]))
  curSample = psList[[taxaLvls[i]]]
  tempOriNTaxa = nrow(otu_table(curSample))
  curSample = phyloseq_filter_prevalence(curSample, prev.trh = 0.1, abund.trh = NULL,
  	threshold_condition = "AND", abund.type = "total")
  ## ALDEx2 with BH
  curALDEx2 <- ALDEx2::aldex(data.frame(phyloseq::otu_table(curSample)), 
                             phyloseq::sample_data(curSample)$diagnosis, 
                             test="t", effect = TRUE, denom="all")
  DA_Summary_Summary_temp = data.frame(
    DAMethod = "ALDEx2",
    TaxaName = rownames(curALDEx2),
    diff_abn = ifelse(curALDEx2$wi.eBH <=0.05, TRUE, FALSE))
  DA_Summary = rbind(DA_Summary, DA_Summary_Summary_temp)
}
```


## Step 5. MaAsLin2
```{r}
outputFileNames = c()
for(i in seq_along(taxaLvls)){
  print(paste0(taxaLvls[i]))
  curSample = psList[[taxaLvls[i]]]
  tempOriNTaxa = nrow(otu_table(curSample))
  curSample = phyloseq_filter_prevalence(curSample, prev.trh = 0.1, abund.trh = NULL,
  	threshold_condition = "AND", abund.type = "total")
  ## MaAsLin2 with BH
  curMaAsLin2 = Maaslin2::Maaslin2(
    input_data = data.frame(phyloseq::otu_table(curSample)), 
    input_metadata = data.frame(phyloseq::sample_data(curSample)), 
    output = paste0(taxaLvls[i]), 
    transform = "AST",
    fixed_effects = "diagnosis",
    normalization = "TSS",
    standardize = FALSE,
    min_prevalence = 0, plot_heatmap = FALSE, plot_scatter = FALSE)
  DA_Summary_Summary_temp = data.frame(
    DAMethod = "MaAsLin2",
    TaxaName = curMaAsLin2$results$feature,
    diff_abn = ifelse(curMaAsLin2$results$qval <=0.05, TRUE, FALSE)
  )
  DA_Summary = rbind(DA_Summary, DA_Summary_Summary_temp)
  ## Delete the temp folder 
  unlink(taxaLvls[i],recursive=TRUE)
}
```

## Step 6. Convert the Differential Abundance Analysis to a Summary
The shiny application accept a data.frame as shown below to be included into the interactive network display. 
```{r}
DASummaryTable = DA_Summary %>%
  pivot_wider(values_from = diff_abn, names_from = DAMethod, id_cols = TaxaName) %>%
  relocate(TaxaName) ## Relocate the TaxaName column to the front <-- Important!
DT::datatable(DASummaryTable, options = list(pageLength = 20))
```

## Step 7. Adding to C3NA Object
Now the DASummaryTable can be easily added to the C3NA object after the comparePhenotypes() function. The Shiny program will use the TaxaName (first column), and the remaining columns as selectable differential abundance methods in TRUE/FALSE format. Please make sure you only has the TaxaName column followed by a number of DA methods (with unique column names)
```{r, eval = FALSE}
## From Tutorial with Demo Step 4:
CancerVsNormal_C3NA = comparePhenotypes(C3NAObj_Comparison =  Cancer_C3NA, 
                                        C3NAObj_Reference = Normal_C3NA,
                                        corCutoff = 0.2, fdr = 0.05, unusefulTaxa = NA, 
                                        nBootstrap = 100, verbose = FALSE)
CancerVsNormal_C3NA$DA = DASummaryTable
```







